Clustering time series based on probability distributions across temporal granularities

### 2021-10-30 main.Rmd
Visual exploration in validation section

## Results

All the methods were fitted to each data designs to check if the formed clusters matches the design, if they are separated well and have low intra-cluster variation. Many of the potentially desired features of clusterings such as separation between clusters, high density within clusters, and distributional shapes can be explored graphically in a more holistic (if subjective) way than by looking at index values. So we use mds and parallel coordinate plots to demonstrate the results. and also for understanding variable importance. Results are summarized below:

- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are. We observe that in all scenarios and for different mean differences, cluster is separated. However, the separation increases with increase in mean differences across scenarios. This is intuitive because with increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.

- Since the designs are constructed in a way that gives importance to three, two and one granularities respectively, we plot the inter cluster differences for each group under the iterations and plot them in a parallel coordinate plot. Figure \ref{fig:parcoord-sim} shows that in the first plot, no variable stands out in the clustering, whereas for the next two designs it is clear that g1, and g1 and g2 have little role to play in the cluster and thus the inter cluster distances are small compared to those variables. For more detailed plots and tables, please refer to the supplementary paper
Each line represents each scenario, hence the point is not to look at cluster separation but more at the fact that which variable led to the clustering. The lines are not colored by group because the goal is not to highlight class division but rather the contribution of the factors to categorization.

 - The performances get worse for js-robust and wpd method for $nT=300$, after which it picks up for larger $nT$ considered in the study. Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example. It is likely that $nT$ would be at least $1000$ with half-hourly data, even if data is only available just for a month. The performance is promising except when the number of observations for a customer is really small.

- For smaller difference between categories, it is expected that method js-nqt would preform better than the other methods.

>>>>

## Visual exploration

All of the approaches were fitted to each data design and for each combination of the considered parameters. The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation. It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values. So we use MDS and parallel coordinate graphs to demonstrate the findings:

- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are. We observe that in all scenarios and for different mean differences, cluster is separated. However, the separation increases with increase in mean differences across scenarios. This is intuitive because with increasing difference between categories, it gets easier for the methods to correctly distinguish the designs. Results corresponding to Scenario (a) is shown here.

- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$. For all simulation settings and scenarios, values are represented as a sequence of lines connected across each axis. One line in the figure, for example, shows the inter-cluster distances for one simulation scenario. The lines are not coloured by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation. The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering. It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (a), two (b), and one (c) significant granularities. Hence, Figure \ref{fig:parcoord-sim} (a), (b), and (c) validate the construction of scenarios (a), (b), and (c) respectively.

 - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study. Although, with the type of residential load data sets, a complete year of load is the minimum requirement to capture predicted differences in winter and summer profiles, for example. Even if the data is only available for a month, $nT$ is expected to be at least $1000$ with half-hourly data. Hence, practically this is not a challenge as long as the performance is promising for higher $nT=300$.

- For smaller difference between categories, it is seen that method js-nqt would preform better js-robust in our study sample. In our study sample, method js-nqt outperforms method js-robust for smaller differences between categories. More testing, however, is required to corroborate this.

For more detailed plots and tables, please refer to the supplementary paper.
